---
title: "Exercise 3"
output: html_document
---

Noon pacific time, 2 hours and 15. (4amAWST)

http://linked.earth/time-uncertain-data-analysis-in-R/correlation.html

```{r}
library(lipdR)
library(geoChronR)
library(ggplot2)
library(magrittr)
library(egg)
```

The Hulu cave speleothem and the GISP2 ice core
```{r}
hulu <- readLipd("https://lipdverse.org/geoChronR-examples/Hulucave.Wang.2001-ens.lpd")
gisp <- readLipd("http://lipdverse.org/geoChronR-examples/GISP2.Alley.2000-ens.lpd")
gisp$paleoData[[1]]$measurementTable[[1]]$year <- NULL
```

```{r}
paleo <- extractTs(hulu) %>% ts2tibble()
paleo$paleoData_variableName
hulu <- mapAgeEnsembleToPaleoData(hulu,
                                  age.var = "age",
                                  model.num = 1,
                                  paleo.depth.var = "depth", 
                                  paleo.meas.table.num = 1)

writeLipd(hulu, "data/Hulucave.Wang.2001.lpd")

hulu.ae <- selectData(hulu,var.name = "ageEnsemble",meas.table.num = 1)
hulu.d18o <- selectData(hulu,var.name = "d18O",meas.table.num = 1)

hulu.plot <- plotTimeseriesEnsRibbons(X = hulu.ae,
                                      Y = hulu.d18o) + ggtitle('Hulu cave')
hulu.plot
```

```{r}
paleo <- extractTs(gisp) %>% ts2tibble()
paleo$paleoData_variableName
gisp <- mapAgeEnsembleToPaleoData(gisp,
                                  age.var = "ageensemble",
                                  model.num = 1,
                                  paleo.depth.var = "age", 
                                  paleo.meas.table.num = 1)

gisp.ae <- selectData(gisp,var.name = "ageEnsemble",meas.table.num = 1)
gisp.temp <- selectData(gisp,var.name = "temp",meas.table.num = 1)

writeLipd(gisp, "data/GISP2.Alley.2000.lpd")

gisp.plot <- plotTimeseriesEnsRibbons(X = gisp.ae,
                                      Y = gisp.temp) + ggtitle('GISP ice core')
gisp.plot
```
They seem to be positively correlated.
```{r}
library(patchwork)
(hulu.plot + xlim(c(50000,35000))) / 
  (gisp.plot + xlim(c(50000,35000)))
```

## 5.3 To corEns() function

```{r}
corout <- corEns(time.1 = gisp.ae,
                 values.1 = gisp.temp,
                 time.2 = hulu.ae,
                 values.2 = hulu.d18o,
                 bin.step = 200,
                 max.ens = 1000,
                 isopersistent  = TRUE,
                 isospectral = TRUE,
                 gaussianize = TRUE)
corout$cor.stats
```

## 5.4 Plotting correlation results
```{r}
raw <- plotCorEns(corout,
           significance.option = "raw",
           use.fdr = FALSE,
           bar.colors = c("grey50", "#E1A81C", "#1B54A8"),
           legend.position = c(0.1, 0.65),
           f.sig.lab.position = c(0.15, 0.61))+
  ggtitle("Distribution of correlation coefficients")

print(raw)
```

Isopersistent is the most similar. eff_n shows the biggest reduction.
```{r fig.height=10}
eff_n <- plotCorEns(corout,
           significance.option = "eff_n",
           use.fdr = FALSE,
           bar.colors = c("grey50", "#E1A81C", "#1B54A8"),
           legend.position = c(0.1, 0.65),
           f.sig.lab.position = c(0.15, 0.61))+
  ggtitle("eff_n")

isopersistent <- plotCorEns(corout,
           significance.option = "isopersistent",
           use.fdr = FALSE,
           bar.colors = c("grey50", "#E1A81C", "#1B54A8"),
           legend.position = c(0.1, 0.65),
           f.sig.lab.position = c(0.15, 0.61))+
  ggtitle("isopersistent")

isospectral <- plotCorEns(corout,
           significance.option = "isospectral",
           use.fdr = FALSE,
           bar.colors = c("grey50", "#E1A81C", "#1B54A8"),
           legend.position = c(0.1, 0.65),
           f.sig.lab.position = c(0.15, 0.61))+
  ggtitle("isospectral")


raw / eff_n / isopersistent/ isospectral
```

### 5.4.2 False-discovery rate testing

```{r}
isoPersistentFdr <- plotCorEns(corout,
           legend.position =c(.85,.8),
           f.sig.lab.position = c(.85,.6),
           significance.option = "isopersistent",
           use.fdr = TRUE)+ggtitle("Isopersistent significance testing with FDR")

print(isoPersistentFdr)
```
## 5.5 Juding signfificance of correlation

GISP has a higher level of self-correlation.

```{r}
corhulu <- corEns(time.1 = hulu.ae,
                 values.1 = hulu.d18o,
                 time.2 = hulu.ae,
                 values.2 = hulu.d18o,
                 bin.step = 200,
                 max.ens = 1000,
                 isopersistent  = TRUE,
                 isospectral = TRUE,
                 gaussianize = TRUE)

isoPersistentFdr_hulu <- plotCorEns(corhulu,
           legend.position =c(.85,.8),
           f.sig.lab.position = c(.85,.6),
           significance.option = "isospectral",
           use.fdr = TRUE)+ ggtitle("Hulu cave isospectral significance testing with FDR")

corgisp <- corEns(time.1 = gisp.ae,
                 values.1 = gisp.temp,
                 time.2 = gisp.ae,
                 values.2 = gisp.temp,
                 bin.step = 200,
                 max.ens = 1000,
                 isopersistent  = TRUE,
                 isospectral = TRUE,
                 gaussianize = TRUE)

isoPersistentFdr_gisp <- plotCorEns(corgisp,
           legend.position =c(.85,.8),
           f.sig.lab.position = c(.85,.6),
           significance.option = "isospectral",
           use.fdr = TRUE)+ggtitle("GISP core isospectral significance testing with FDR")
isoPersistentFdr_hulu / isoPersistentFdr_gisp
```

## 5.6 Chapter Project

Todo